---
title: "Week1"
author: "Ishaan Das"
date: "October 30, 2018"
output: html_document
---

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(lattice)
```

###Let us start by reading in the IRIS Dataset

```{r}
str(iris)
```

###Next we can check the first 6 and last 6 rows using the head and tail function to check for consistency of observations. We can also check for missing values.

```{r}
head(iris)
#Here we see the first 5 rows of the dataset and all columns have values.
#Next we can execute the tail function.
tail(iris)
#The output shows us the last 5 rows of the dataset and all columns have values.
#Next we execute a function that will provide us with the number of null values.
colSums(is.na(iris))
#This function shows us all the null values and we can see that there are 0 null values for all 5 variables. 
```

###Next we can conduct a summary of the data by executing various functions.

```{r}
#We can start by checking for the ranges of each variable to check if any of the dimensions are 0 because length and width can't be 0. And if our dataset does contain 0 then it would indicate a potential dataset error.
range(iris$Sepal.Length)
range(iris$Sepal.Width)
range(iris$Petal.Length)
range(iris$Petal.Width)
#None of our observations are 0 which indicates a good dataset.

#Next we can check how many of our observations lie in each flower species
table(iris$Species)
#the output indicates an equal distribution across all three flower species
```

###Given equal number of observations we can create some plots.

```{r}
#First let us create some box plots to see the range of the observations and if there are any outliers.
boxplot(iris[,c(1:4)], ylab = "Measurements in centimeters")
#This boxplot shows us the range of each variable.

#We are interested in knowing how sepal length differs for each species.
boxplot(iris$Sepal.Length ~ iris$Species, ulab = "Measurements in centimeters")
#Setosas have the smallest sepal lengths followed by versicolor and then by virginica.

#We are also interested in knowing how sepal width differs for each species.
boxplot(iris$Sepal.Width ~ iris$Species, ulab = "Measurements in centimeters")
#For sepal width range is different for each species.

#We can check sepal length histograms for each species.
histogram( ~ Sepal.Length | Species, data = iris)

#We can check sepal width histograms for each species.
histogram( ~ Sepal.Width | Species, data = iris)

#We can check petal length histograms for each species.
histogram( ~ Petal.Length | Species, data = iris)

#We can check sepal width histograms for each species.
histogram( ~ Petal.Width | Species, data = iris)
```

##There seems to be some similarities in the petal histograms indicating a possible relationship between petal.length and petal.width.